---
title: "Stats 506, F20, Problem Set 5"
author: "Zhihao Xu, xuzhihao@umich.edu"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = FALSE)
source('./ps5_q1.R')
```

## About
This is the write-up solution to Problem Set 5 for Stats 506 in Fall 2020.

### Code Organization

## Question 1


> For which Census Division has most air conditioning equipment in 2009 and 2015 and which change most from 2009 to 2015? Is this difference larger in urban or rural areas?

Click right to see the R code for answering this question. 
```{r codeq1, results='asis'}
cat('```r\n')
writeLines(readLines('./ps5_q1.R'))
cat('```')
```

### Tables{.tabset .tabset-pills .tabset-fade}
#### Proportion in 2009
```{r tab09}
DT::datatable(tab09, escape = FALSE)
```

#### Proportion in 2015
```{r tab15}
DT::datatable(tab15, escape = FALSE)
```

#### Proportion changes from 2009 to 2015
```{r tabdif}
DT::datatable(tabdif, escape = FALSE)
```

### Figures{.tabset .tabset-pills .tabset-fade}
#### Proportion in 2009
```{r fig09,fig.cap=cap1,fig.width=8}
p1
```

#### Proportion in 2015
```{r fig15,fig.cap=cap2,fig.width=8}
p2
```

#### Proportion changes from 2009 to 2015
```{r figdif,fig.cap=cap3,fig.width=8}
p3
```


