---
title: "Stats 506, F20, Problem Set 3"
author: "Zhihao Xu, xuzhihao@umich.edu"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = FALSE)
options(knitr.kable.NA = '')
```


```{r libraries, include=TRUE, message = FALSE}
library(tidyverse)
```

```{r directory, include=TRUE, message = FALSE}
path = './output/'
```
## About
This is the write-up solution to Problem Set 3 for Stats 506 in Fall 2020.

### Code Organization
- `ps3_q1.do` consists of the `stata` code used in Question 1. Please comment the `mkdir` line after the first run.
- `ps3_q2.do` consists of the `stata` code used in Question 2
- `ps3_q3.do` consists of the `stata` code used in Question 3. Please comment the `mkdir` line after the first run.
- `output` folder consists all the tables exported from `stata`, and the sub-folder `ohx` consists all the `csv` files exported from `stata` for Question 3. 
- `ps3_writeup.R` imports all the data exported from `stata` stored in the `output` folder and builds the page your are reading. Code for figures and tables can be found here.


## Question 1

Here we construct the balance tables for each of the demographics required. We use chi-squared test for marginal associations between categorical variables and whether the dentition exam is complete, and t-test for continuous variable. For continuous variable, present `mean(IQR)` and for categorical variables, present `n(%)`.
```{r read_data, warning = FALSE}
sheets = c("edu20p", "gender20m", "gender20p", "race20m", "race20p")
sheetspath = paste0(path, sheets)
result_list = vector(mode = 'list')
for (sheet in sheetspath){
  file = paste0(sheet, ".csv")
  result_list[[sheet]] = read_delim( file, delim = ',' )
}
excel_sheets = c("20m", "20p")
for (sheet in excel_sheets) {
 result_list[[sheet]] =   readxl::read_xlsx(paste0(path,"age.xlsx"), 
                                            sheet = sheet, col_names = F)
}
names(result_list) = c(sheets, "age20m", "age20p")
```



### Age Under 20
### {.tabset .tabset-pills .tabset-fade}
#### Age
```{r tabage20m}
result_list$age20m[-4] %>%
  transmute(
    ohx = ...1,
    mean_IQR = paste0(sprintf("%4.2f", ...2)," (",
                 sprintf("%2.0f", ...3),")")) %>%
  bind_rows(tribble(~ohx, ~mean_IQR,
        "$T$", 
        paste0(sprintf("%4.2f", result_list$age20m[1,4])," (p=", 
               sprintf("%3.2f", result_list$age20m[2,4]),")"))) %>%
  knitr::kable(format = "html", 
               col.names = c('Ohx Status', 'Age')) %>%
  kableExtra::kable_styling('striped', full_width = TRUE) %>%
  kableExtra::group_rows("Test Marginal Associations", 
                         start_row = 3, end_row = 3) 
```

#### Gender
```{r tabgender20m}
chi2 = result_list$gender20m[9,2]
p = as.numeric(result_list$gender20m[10,2])
result_list$gender20m[c(3,5,7),] %>%
  bind_cols(result_list$gender20m[c(4,6,8),2:4]) %>%
  transmute(
    ohx = X1,
    male = paste0(format(as.numeric(....2))," (",
                 sprintf("%4.2f", as.numeric(....5)),")"),
    female = paste0(format(as.numeric(X3...3))," (",
                 sprintf("%4.2f", as.numeric(X3...6)),")"),
    total = paste0(format(as.numeric(X4...4))," (",
                 sprintf("%4.2f", as.numeric(X4...7)),")")
  ) %>%
  bind_rows(tribble(~ohx, ~male,
        "$\\chi^2$", 
        paste0(sprintf("%4.2f", chi2)," (p=", 
               sprintf("%3.2f", p),")"))) %>%
  knitr::kable(format = "html", 
               col.names = c('Ohx Status', 'Male', 'Female', 'Tatal')) %>%
  kableExtra::kable_styling('striped', full_width = TRUE) %>%
  kableExtra::group_rows("Test Marginal Associations", 
                         start_row = 4, end_row = 4) 
```
#### Race/Ethnicity
```{r tabrace20m}
chi2 = result_list$race20m[9,2]
p = as.numeric(result_list$race20m[10,2])
result_list$race20m[c(3,5,7),] %>%
  bind_cols(result_list$race20m[c(4,6,8),2:8]) %>%
  transmute(
    ohx = X1,
    ma = paste0(format(as.numeric(....2))," (",
                 sprintf("%4.2f", as.numeric(....9)),")"),
    oh = paste0(format(as.numeric(X3...3))," (",
                 sprintf("%4.2f", as.numeric(X3...10)),")"),
    nhw = paste0(format(as.numeric(X4...4))," (",
                 sprintf("%4.2f", as.numeric(X4...11)),")"),
    nhb = paste0(format(as.numeric(X5...5))," (",
                 sprintf("%4.2f", as.numeric(X5...12)),")"),
    nha = paste0(format(as.numeric(X6...6))," (",
                 sprintf("%4.2f", as.numeric(X6...13)),")"),
    or = paste0(format(as.numeric(X7...7))," (",
                 sprintf("%4.2f", as.numeric(X7...14)),")"),
    total = paste0(format(as.numeric(X8...8))," (",
                 sprintf("%4.2f", as.numeric(X8...15)),")")
  ) %>%
  bind_rows(tribble(~ohx, ~ma,
        "$\\chi^2$", 
        paste0(sprintf("%4.2f", chi2)," (p=", sprintf("%3.2f", p),")"))) %>%
  knitr::kable(format = "html", 
               col.names = c('Ohx Status', 'Mexican American', 
                             'Other Hispanic', 'Non-Hispanic White', 
                             'Non-Hispanic Black', 'Non-Hispanic Asian', 
                             'Other Race','Tatal'))%>%
  kableExtra::kable_styling('striped', full_width = TRUE) %>%
  kableExtra::group_rows("Test Marginal Associations", 
                         start_row = 4, end_row = 4)
```




### Age Over 20
### {.tabset .tabset-pills .tabset-fade}
#### Age
```{r tabage20p}
result_list$age20p[-4] %>%
  transmute(
    ohx = ...1,
    mean_IQR = paste0(sprintf("%4.2f", ...2)," (",
                 sprintf("%2.0f", ...3),")")) %>%
  bind_rows(tribble(~ohx, ~mean_IQR,
        "$T$", 
        paste0(sprintf("%4.2f", result_list$age20p[1,4])," (p=", 
               sprintf("%3.2f", result_list$age20p[2,4]),")")))%>%
  knitr::kable(format = "html", 
               col.names = c('Ohx Status', 'Age')) %>%
  kableExtra::kable_styling('striped', full_width = TRUE) %>%
  kableExtra::group_rows("Test Marginal Associations", 
                         start_row = 3, end_row = 3) 
```

#### Gender
```{r tabgender20p}
chi2 = result_list$gender20p[9,2]
p = as.numeric(result_list$gender20p[10,2])
result_list$gender20p[c(3,5,7),] %>%
  bind_cols(result_list$gender20p[c(4,6,8),2:4]) %>%
  transmute(
    ohx = X1,
    male = paste0(format(as.numeric(....2))," (",
                 sprintf("%4.2f", as.numeric(....5)),")"),
    female = paste0(format(as.numeric(X3...3))," (",
                 sprintf("%4.2f", as.numeric(X3...6)),")"),
    total = paste0(format(as.numeric(X4...4))," (",
                 sprintf("%4.2f", as.numeric(X4...7)),")")
  ) %>%
  bind_rows(tribble(~ohx, ~male,
        "$\\chi^2$", 
        paste0(sprintf("%4.2f", chi2)," (p=", 
               sprintf("%3.2f", p),")"))) %>%
  knitr::kable(format = "html",
               col.names = c('Ohx Status', 'Male', 'Female', 'Tatal')) %>%
  kableExtra::kable_styling('striped', full_width = TRUE) %>%
  kableExtra::group_rows("Test Marginal Associations", 
                         start_row = 4, end_row = 4) 
```

#### Race/Ethnicity
```{r tabrace20p}
chi2 = result_list$race20p[9,2]
p = as.numeric(result_list$race20p[10,2])
result_list$race20p[c(3,5,7),] %>%
  bind_cols(result_list$race20p[c(4,6,8),2:8]) %>%
  transmute(
    ohx = X1,
    ma = paste0(format(as.numeric(....2))," (",
                 sprintf("%4.2f", as.numeric(....9)),")"),
    oh = paste0(format(as.numeric(X3...3))," (",
                 sprintf("%4.2f", as.numeric(X3...10)),")"),
    nhw = paste0(format(as.numeric(X4...4))," (",
                 sprintf("%4.2f", as.numeric(X4...11)),")"),
    nhb = paste0(format(as.numeric(X5...5))," (",
                 sprintf("%4.2f", as.numeric(X5...12)),")"),
    nha = paste0(format(as.numeric(X6...6))," (",
                 sprintf("%4.2f", as.numeric(X6...13)),")"),
    or = paste0(format(as.numeric(X7...7))," (",
                 sprintf("%4.2f", as.numeric(X7...14)),")"),
    total = paste0(format(as.numeric(X8...8))," (",
                 sprintf("%4.2f", as.numeric(X8...15)),")")
  ) %>%
  bind_rows(tribble(~ohx, ~ma,
        "$\\chi^2$", 
        paste0(sprintf("%4.2f", chi2)," (p=", sprintf("%3.2f", p),")"))) %>%
  knitr::kable(format = "html", 
               col.names = c('Ohx Status', 'Mexican American', 
                             'Other Hispanic', 'Non-Hispanic White', 
                             'Non-Hispanic Black', 'Non-Hispanic Asian', 
                             'Other Race','Tatal'))%>%
  kableExtra::kable_styling('striped', full_width = TRUE) %>%
  kableExtra::group_rows("Test Marginal Associations",
                         start_row = 4, end_row = 4)
```

#### Education Level
```{r tabedu20p}
chi2 = result_list$edu20p[9,2]
p = as.numeric(result_list$edu20p[10,2])
result_list$edu20p[c(3,5,7),] %>%
  bind_cols(result_list$edu20p[c(4,6,8),2:4]) %>%
  transmute(
    ohx = X1,
    sc = paste0(format(as.numeric(....2))," (",
                 sprintf("%4.2f", as.numeric(....5)),")"),
    other = paste0(format(as.numeric(X3...3))," (",
                 sprintf("%4.2f", as.numeric(X3...6)),")"),
    total = paste0(format(as.numeric(X4...4))," (",
                 sprintf("%4.2f", as.numeric(X4...7)),")")
  ) %>%
  bind_rows(tribble(~ohx, ~sc,
        "$\\chi^2$", 
        paste0(sprintf("%4.2f", chi2)," (p=", sprintf("%3.2f", p),")"))) %>%
  knitr::kable(format = "html", 
               col.names = c('Ohx Status', 'Some college/College graduate', 
                             'No college/Unknown', 'Tatal')) %>%
  kableExtra::kable_styling('striped', full_width = TRUE) %>%
  kableExtra::group_rows("Test Marginal Associations", 
                         start_row = 4, end_row = 4) 
```






## Question 2

In this question, we tried to fit the logistic regression model about whether an NHANES participant has a complete dentition exam conditional on the demographics include in the Question 1. We tried some different models like using quadratic or cubic function of `age`, and considering interactions between each predictor. The final model we choose is `c.age##c.age##i.u20 i.riagendr i.n_edu` with AIC `14951.27`.

```{r q2read_data, warning = FALSE}
result_list = vector(mode = 'list')
excel_sheets = c("logitreg", "regout")
for (sheet in excel_sheets) {
 result_list[[sheet]] =   readxl::read_xlsx(paste0(path,"ps3_q2.xlsx"), 
                                            sheet = sheet, col_names = T)
}
```

### {.tabset .tabset-pills .tabset-fade}
#### Regression Table of Final Model
```{r q2model}
result_list$regout %>%
  filter(is.na(ohx)==F) %>%
  knitr::kable(format = "html", 
               col.names = c('Parameter', 'Coef.',  'Std. Err.', 
                             'z', 'P>|z|', 'lwr', 'upr')) %>%
  kableExtra::kable_styling('striped', full_width = TRUE)
```

#### AIC
```{r q2aic}
result_list$logitreg %>%
  knitr::kable(format = "html", 
               col.names = c('Model', 'AIC')) %>%
  kableExtra::kable_styling('striped', full_width = TRUE)
```


## Question 3

In this question, we tried to investigate the relationship between each teeth’s status and age, and quantify how teeth progress as we age. Here we construct a big figure to present our findings. Through the following figure, we can see that when age is under 6, the proportion of `Primary tooth (delicious) present` and `Tooth not present` is relatively high. And, these two proportions are decreasing sharply as age increases, which is decreased to around 0 when age is greater than 12. For the proportion of `Permanent tooth present`, it is close to 0 when age is less than 6, and increases when age is from 6 to 12 depends on the teeth number. After the age is greater than 18, this proportion starts to slightly decrease as age increases. The proportion of `Dental implant` and `Permanent dental root fragment` are very close to 0 when age is less than 18, and increase after that. Compared to other proportions, the confidence interval of these two proportions are relatively large, especially for `Dental implant`, which indicates that the diversity of `Dental implant` level is quite large compared with other levels.


```{r q3readdata, warning=FALSE}
idx = sprintf("%02d", 1:32)
sheetspath = paste0(path, "ohx/ohx",idx,"tc.csv")
result_list = vector(mode = 'list')
for (sheet in sheetspath){
  result_list[[sheet]] = read_delim( sheet, delim = ',', skip=3, 
                                     col_names = T, skip_empty_rows = T, 
                                     col_types = "cnnnnnn")
}
names(result_list) = paste0('ohx',idx)
result_list$ohx03 = result_list$ohx03[,-2]
```

```{r q3dataprep}
change_name = function(name){
  n = floor(length(name)/2)
  col = name[2:(2+n-1)]
  col = sapply(col, substr, 3, 3)
  col = c("age",paste0("l",col), paste0("s",col))
  return(col)
}

full_data = tibble(
  age = 0,
  teeth = 0,
  level = "1",
  prop = 0,
  se = 0
)
for (ohx in names(result_list)){
  ncol = dim(result_list[[`ohx`]])[2]-2
  result_list[[`ohx`]] = result_list[[`ohx`]][seq(1,27,2),1:(2+ncol-1)] %>%
    bind_cols(result_list[[`ohx`]][seq(2,28,2),2:(2+ncol-1)]) %>%
    rename_with(change_name) %>%
    mutate(age = 1:14)
  num = as.numeric(substr(ohx,4,5))
  result_list[[`ohx`]] = result_list[[`ohx`]]  %>% 
    pivot_longer(
      cols = !1,
      names_pattern = "(l|s)([[:digit:]])",
      names_to = c(".value", "level")
    )  %>%
    transmute(
      age = age,
      teeth = num,
      level = level,
      prop = l,
      se = s
    )
  full_data = full_data %>%
    bind_rows(result_list[[`ohx`]])
}
m = qnorm(.975)
i = seq(0,80,6)
age_labels = paste0("[",i,",",i+6,")")
teeth_labels = c("Primary tooth (deciduous) present",
                 "Permanent tooth present",
                 "Dental implant",
                 "Tooth not present",
                 "Permanent dental root fragment present")
full_data = full_data[-1,] %>%
  mutate(
    age = factor(age, labels = age_labels),
    teeth = factor(teeth),
    level = factor(level, labels = teeth_labels),
    lwr = pmax(prop - m * se,0),
    upr = pmin(prop + m * se,1)
  ) %>% select(-se)
```



```{r fig.height=50, fig.width=12, fig.align="center", fig.cap=cap}
cap = paste(
  "**Figure 1.** Porportion of Each Teeth's Status Level ",
  "by Age"
)
full_data%>%
  group_by(teeth, level) %>%
  ggplot(aes(x = prop, y = teeth, shape = level)) +
    geom_point(aes(color = level), position = position_dodge(0.3)) +
    geom_errorbarh( aes(xmin = lwr, xmax = upr, color=level), 
                    position = position_dodge(0.3)) + 
    xlab('Proportion of Each Level') +
    ylab('Teeth Number from 1 to 32') +
    facet_wrap(vars(age),nrow = 7, ncol=2) + 
    theme_bw() + 
    theme(legend.position="top")
```

