---
title: "Stats 506, F20, Problem Set 1"
author: "Zhihao Xu, xuzhihao@umich.edu"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---

In this assignment, `ps1_q1_ohxden.sh` is the bash file for Question 1 Part A, 
`ps1_q1_demo.sh` is the bash file for Part B, `xpt2csv.R` is the Rcode used to 
transfer the XPT file to csv file, and `ps1_q2.R` is the R code for Question 2.

```{r setup, include=FALSE}
# 79: -------------------------------------------------------------------------
#! We generally don't need to see the code in the results document.
knitr::opts_chunk$set(echo = FALSE)
#! Make sure all chunks have a label. This one is labelled 'setup'.
```

## Question 1

### Part A
Here is the solution text for `ps1_q1_ohxden.sh`. In the resulting datasets, 
there are totally 35909 observations and 62 variables.

```{r ps1_q1a, results = 'asis'}
cat('```bash\n')
writeLines(readLines('./ps1_q1_ohxden.sh'))
cat('```')
```


### Part B
Here is the solution text for `ps1_q1_demo.sh`. In the resulting datasets, 
there are totally 39156 observations and 10 variables.

```{r ps1_q1b, results = 'asis'}
cat('```bash\n')
writeLines(readLines('./ps1_q1_demo.sh'))
cat('```')
```

\pagebreak

## Question 2

### Part A and B
```{r ps1_q2ab, results = 'asis'}
cat('```r\n')
writeLines(readLines('./ps1_q2.R'))
cat('```')
```

\pagebreak
### Part C
```{r rocrf, include=FALSE}
source("ps1_q2.R")
roc = perf_roc(data[,2],data[,1],"none")
pr = perf_pr(data[,2],data[,1],"none")
```

The AUC-ROC is `r roc$area_roc` and the AUC-PR is `r pr$area_pr`. 
Attached are both the base R and ggplot2 versions of plots showing the curves.

```{r q2_roc1, echo=FALSE, fig.height=4, fig.width=5.5,fig.cap=cap}
roc1 = perf_roc(data[,2],data[,1],"base")
cap = paste0("ROC Curve by base R Graphics")
```

```{r q2_roc2, echo=FALSE, fig.height=3, fig.width=5,fig.cap=cap}
roc2 = perf_roc(data[,2],data[,1],"ggplot2")
cap = paste0("ROC Curve by ggplot2")
```

```{r q2_pr1, echo=FALSE, fig.height=4, fig.width=5.5,fig.cap=cap}
pr1 = perf_pr(data[,2],data[,1],"base")
cap = paste0("Prediction-Recall Curve by base R Graphics")
```

```{r q2_pr2, echo=FALSE, fig.height=3, fig.width=5,fig.cap=cap}
pr2 = perf_pr(data[,2],data[,1],"ggplot2")
cap = paste0("Prediction-Recall Curve by ggplot2")
```

