---
title: "Stats 506, F20, Problem Set 4"
author: "Zhihao Xu, xuzhihao@umich.edu"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = FALSE)
options(knitr.kable.NA = '')
```


```{r libraries, include=TRUE, message = FALSE}
library(tidyverse)
```

```{r directory, include=TRUE, message = FALSE}
path = './output/'
```
## About
This is the write-up solution to Problem Set 3 for Stats 506 in Fall 2020.

### Code Organization
- `ps4_q1.sas` consists of the `SAS` code used in Question 1. You need to create the `output/` folder before run. My `SAS` version is University Edition, the paths in the code are different from the path on the computer, which is start from `/folders/myfolders`. You need to change the path if it is running on other computer.
- `ps4_q2.sas` consists of the `SAS` code used in Question 2.
- `output` folder consists all the tables exported from `SAS`, which will be used in this write-up file for display.
- `ps4_writeup.Rmd` builds the page you are reading. Code for figures and tables can be found here.

## Question 1

<details>
<summary> Click to view `SAS code for Question 1`. </summary>
```{r}
writeLines(readLines('./ps4_q1.sas'))
```
</details>
### 
(a) Compare the average number of televisions (`TVCOLOR`) in US homes in 2009 and 2015 by Census Region

```{r q1a_readdata}
sheets = c("ps4_q1_a09", "ps4_q1_a15", "ps4_q1_aii")
sheetspath = paste0(path, sheets)
result_q1a = vector(mode = 'list')
for (sheet in sheetspath){
  file = paste0(sheet, ".csv")
  result_q1a[[sheet]] = read_delim( file, delim = ',' )
}
```

### {.tabset .tabset-pills .tabset-fade}
#### TV Number 2009
```{r q1ai09, fig.height=5, fig.width=8, fig.cap=cap, fig.align="center"}
cap = paste(
  "**Figure 1.** Average Number of TV in US Homes in 2009"
)
region = as_tibble_col(c("Northeast Census Region",
                         "Midwest Census Region",
                         "South Census Region",
                         "West Census Region"), 
                       column_name = "REGIONC")
result_q1a$`./output/ps4_q1_a09`[,1] = region
result_q1a$`./output/ps4_q1_a09` %>%
  ggplot(aes(x = avg_tv09, y = REGIONC)) +
  geom_point() +
  geom_errorbarh( aes(xmin = lwr, xmax = upr)) + 
  xlab('Mean Number of Televisions') +
  theme_bw()
```

#### TV Number 2015
```{r q1ai15, fig.height=5, fig.width=8, fig.cap=cap, fig.align="center"}
cap = paste(
  "**Figure 2.** Average Number of TV in US Homes in 2015"
)
result_q1a$`./output/ps4_q1_a15`[,1] = region
result_q1a$`./output/ps4_q1_a15` %>%
  ggplot(aes(x = avg_tv15, y = REGIONC)) +
  geom_point() +
  geom_errorbarh( aes(xmin = lwr, xmax = upr)) + 
  xlab('Mean Number of Televisions') +
  theme_bw()
```

#### Difference of TV Number from 2009 to 2015
```{r q1aii, fig.height=5, fig.width=8, fig.cap=cap, fig.align="center"}
cap = paste(
  "**Figure 3.** Difference of Average Number of TV in US Homes ",
  "from 2009 to 2015"
)
result_q1a$`./output/ps4_q1_a15`[,1] %>%
  bind_cols(result_q1a$`./output/ps4_q1_aii`) %>%
  ggplot(aes(x = dif, y = REGIONC)) +
  geom_point() +
  geom_errorbarh( aes(xmin = lwr, xmax = upr)) + 
  xlab('Mean of the Number of Televisions') +
  theme_bw()
```

#### Estimate and Difference in Table
```{r q1aiii}
cap = paste(
  "**Table 1.** Estimate Value of Average TV Number in US Homes ",
  "for Both 2009 and 2015, and Their Difference"
)
result_q1a$`./output/ps4_q1_a09`[,1] %>%
  bind_cols(
    result_q1a$`./output/ps4_q1_a09` %>%
      transmute(
        tv09 = paste0(sprintf("%4.2f", avg_tv09)," (",
                     sprintf("%4.2f", lwr),", ",
                     sprintf("%4.2f", upr),")"),
      )
  ) %>%
  bind_cols(
    result_q1a$`./output/ps4_q1_a15` %>%
      transmute(
        tv15 = paste0(sprintf("%4.2f", avg_tv15)," (",
                     sprintf("%4.2f", lwr),", ",
                     sprintf("%4.2f", upr),")"),
      )
  ) %>%
  bind_cols(
    result_q1a$`./output/ps4_q1_aii` %>%
      transmute(
        dif = paste0(sprintf("%4.2f", dif)," (",
                     sprintf("%4.2f", lwr),", ",
                     sprintf("%4.2f", upr),")"),
      )
  )%>%
  knitr::kable(format = "html", caption = cap,
               col.names = c('Census Region', 'TV Number 2009',
                             'TV Number 2015', 'Difference')) %>%
  kableExtra::kable_styling('striped', full_width = TRUE)
```


### 
(b) Compare the proportion of primary televisions by display type for most used television (`TVTYPE1`) in US homes in 2009 and 2015 by Census Region

### {.tabset .tabset-pills .tabset-fade}

```{r q1b_readdata}
sheets = c("ps4_q1_b09", "ps4_q1_b15", "ps4_q1_bii")
sheetspath = paste0(path, sheets)
result_q1b = vector(mode = 'list')
for (sheet in sheetspath){
  file = paste0(sheet, ".csv")
  result_q1b[[sheet]] = read_delim( file, delim = ',' )
}
```

#### TV Type 2009
```{r q1bi09, fig.height=6, fig.width=9, fig.cap=cap, fig.align="center"}
cap = paste(
  "**Figure 4.** Proportion of Most Used TV in US Homes in 2009"
)
region = as_tibble_col(rep(c("Northeast Census Region",
                         "Midwest Census Region",
                         "South Census Region",
                         "West Census Region"), each = 6), 
                       column_name = "REGIONC")
tvtype = as_tibble_col(rep(c("Not Applicable",
                         "Standard Tube", "LCD", "Plasma",
                         "Projection", "LED"), 4), 
                       column_name = "TVTYPE1")
result_q1b$`./output/ps4_q1_b09`[,1] = region
result_q1b$`./output/ps4_q1_b09`[,2] = tvtype
result_q1b$`./output/ps4_q1_b09` %>%
  group_by(REGIONC, TVTYPE1) %>%
  ggplot(aes(x = RowPercent, y = REGIONC, shape=TVTYPE1)) +
  geom_point(aes(color = TVTYPE1), position = position_dodge(0.3)) +
  geom_errorbarh( aes(xmin = lwr, xmax = upr, color=TVTYPE1), 
                position = position_dodge(0.3)) + 
  xlab('Proportion of TV Types') +
  theme_bw() + 
  theme(legend.position="top")
```

#### TV Type 2015
```{r q1bi15, fig.height=6, fig.width=9, fig.cap=cap, fig.align="center"}
cap = paste(
  "**Figure 5.** Proportion of Most Used TV in US Homes in 2015"
)
result_q1b$`./output/ps4_q1_b15`[,1] = region
result_q1b$`./output/ps4_q1_b15`[,2] = tvtype
result_q1b$`./output/ps4_q1_b15` %>%
  group_by(REGIONC, TVTYPE1) %>%
  ggplot(aes(x = RowPercent, y = REGIONC, shape=TVTYPE1)) +
  geom_point(aes(color = TVTYPE1), position = position_dodge(0.3)) +
  geom_errorbarh( aes(xmin = lwr, xmax = upr, color=TVTYPE1), 
                position = position_dodge(0.3)) + 
  xlab('Proportion of TV Types') +
  theme_bw() + 
  theme(legend.position="top")
```

#### Difference of TV Type from 2009 to 2015
```{r q1bii, fig.height=6, fig.width=9, fig.cap=cap, fig.align="center"}
cap = paste(
  "**Figure 6.** Difference in Proportion of Most Used TV in US Homes ", 
  "from 2009 to 2015"
)
result_q1b$`./output/ps4_q1_bii`[,1] = tvtype
region %>%
  bind_cols(result_q1b$`./output/ps4_q1_bii`) %>%
  group_by(REGIONC, TVTYPE1) %>%
  ggplot(aes(x = dif, y = REGIONC, shape=TVTYPE1)) +
  geom_point(aes(color = TVTYPE1), position = position_dodge(0.3)) +
  geom_errorbarh( aes(xmin = lwr, xmax = upr, color=TVTYPE1), 
                position = position_dodge(0.3)) + 
  xlab('Proportion of TV Types') +
  theme_bw() + 
  theme(legend.position="top")
```


#### Estimate and Difference in Table
```{r q1biii}
cap = paste(
  "**Table 2.** Estimate Value of the Proportion of Most Used TV in US Homes ",
  "for Both 2009 and 2015, and Their Difference"
)
region %>%
  bind_cols(tvtype) %>%
  bind_cols(
    result_q1b$`./output/ps4_q1_b09` %>%
      transmute(
        tv09 = paste0(sprintf("%5.3f", RowPercent/100)," (",
                     sprintf("%5.3f", lwr/100),", ",
                     sprintf("%5.3f", upr/100),")"),
      )
  ) %>%
  bind_cols(
    result_q1b$`./output/ps4_q1_b15` %>%
      transmute(
        tv09 = paste0(sprintf("%5.3f", RowPercent/100)," (",
                     sprintf("%5.3f", lwr/100),", ",
                     sprintf("%5.3f", upr/100),")"),
      )
  )%>%
  bind_cols(
    result_q1b$`./output/ps4_q1_bii` %>%
      transmute(
        tv09 = paste0(sprintf("%5.3f", dif/100)," (",
                     sprintf("%5.3f", lwr/100),", ",
                     sprintf("%5.3f", upr/100),")"),
      )
  )%>%
  knitr::kable(format = "html", caption = cap,
               col.names = c('Census Region', 'TV Type', 'Proportion 2009', 
                             'Proportion 2015', 'Difference')) %>%
  kableExtra::kable_styling('striped', full_width = TRUE)

```

## Question 2
<details>
<summary> Click to view `SAS code for Question 2`. </summary>
```{r}
writeLines(readLines('./ps4_q2.sas'))
```
</details>

Here I just pick the first tooth `OHX01TC` as response.

```{r q2a}
sheets = c("ps4_q2_1", "ps4_q2_2", "ps4_q2_3", "ps4_q2_4")
sheetspath = paste0(path, sheets)
result_q2 = vector(mode = 'list')
for (sheet in sheetspath){
  file = paste0(sheet, ".csv")
  result_q2[[sheet]] = read_delim( file, delim = ',' )
}
```

<html>
<table style="margin-left: auto; margin-right: auto;" class="table table-bordered table-striped table-condensed">>

<tr>
<th> Logistics Regression without weight `proc logistic`</th>
<th> Logistics Regression with weight `proc surveylogistic`</th>
</tr>

<tr>
<td> Analysis of Maximum Likelihood Estimates </td>
<td> Analysis of Maximum Likelihood Estimates </td>
</tr>

<tr>
<td>
```{r q2_1}
result_q2$`./output/ps4_q2_1` %>%
  select(-`_ESTTYPE_`) %>%
  mutate(
    Estimate = sprintf("%4.2f", Estimate),
    StdErr = sprintf("%4.2f", StdErr),
    WaldChiSq = sprintf("%4.2f", WaldChiSq)
  ) %>%
  knitr::kable(format = "html") %>%
  kableExtra::kable_styling('striped', full_width = TRUE)
```
</td>
<td> 
```{r q2_2}
result_q2$`./output/ps4_q2_3` %>%
  select(-tValue, -ProbT) %>%
  mutate(
    Estimate = sprintf("%4.2f", Estimate),
    StdErr = sprintf("%4.2f", StdErr),
    WaldChiSq = sprintf("%4.2f", WaldChiSq)
  ) %>%
  knitr::kable(format = "html") %>%
  kableExtra::kable_styling('striped', full_width = TRUE)
```
</td>
</tr>

<tr>
<td> Model Fit Statistics </td>
<td> Model Fit Statistics </td>
</tr>

<tr>
<td>
```{r q2_3}
result_q2$`./output/ps4_q2_2` %>%
  knitr::kable(format = "html") %>%
  kableExtra::kable_styling('striped', full_width = TRUE)
```
</td>
<td> 
```{r q2_4}
result_q2$`./output/ps4_q2_4` %>%
  knitr::kable(format = "html") %>%
  kableExtra::kable_styling('striped', full_width = TRUE)
```
</td>
</tr>
</table>
</html>








